Dependent Object Types

### 2012-08-06 article.tex
defs of logical relations
fix indices
rel s in addition to g, do not overload, acks todo
logical rels

\usepackage{amsmath}

@@

@@

@@

\todo{complete proof!}

@@

%\appendix
%\section{Appendix Title}
%\acks

>>>>

\usepackage{amsmath}
\usepackage{amssymb}

@@

\usepackage{stmaryrd}

@@

% Logical relations
\newcommand{\relv}[4]{\mathcal{V}_{#1;#2;#3}\llbracket#4\rrbracket}
\newcommand{\rele}[4]{\mathcal{E}_{#1;#2;#3}\llbracket#4\rrbracket}
\newcommand{\rels}[3]{\mathcal{S}_{#1;#2}\llbracket#3\rrbracket}
\newcommand{\relg}[3]{\mathcal{G}_{#1;#2}\llbracket#3\rrbracket}
\newcommand{\irred}[2]{\text{irred }(#1,#2)}
\newcommand{\andl}{\;\wedge}
\newcommand{\orl}{\vee}
\newcommand{\impliesl}{\rightarrow}
\newcommand{\reductionl}[5]{#1 \operatorname{|} #2 \;\rightarrow^{#5}\; #3 \operatorname{|} #4}
\newcommand{\ds}{\,\vDash\,}

@@

\subsection{Type Safety}
Type-safety states that a well-typed program doesn't get stuck. More
formally: If $- \ts t \typ T$ and $\reductionl t - {t'} {s'} *$ then
either $t'$ is a value or $\exists t'', s'', \reduction {t'} {s'}
{t''} {s''}$.

Our strategy is to define a logical relation $\Gamma \ds t : T$, such
that $\Gamma \ts t : T$ implies $\Gamma \ds t : T$ implies
type-safety.

\subsection{Logical Relations}

\subsubsection{$\mathcal{V}$}
$\relv k \Gamma s T$ defines the set of values that appear to have
type $T$ when taking at most $k$ steps. $\Gamma$ and $s$ must agree:
$\dom(\Gamma) = \dom(s)$ and $\forall (x : T) \in \Gamma, x \in \relv
k \Gamma s T$. A variable $y$ belongs to $\relv 0 \Gamma s T$ simply
by being in the store. In addition, it belongs to $\relv k \Gamma s T$
for $k > 0$, if it defines all method and value labels in the
expansion of $T$ appropriately for $j < k$ steps.
\begin{align*}
&\relv k \Gamma s T = \{ y \ |\\
&\gap \Gamma \ts T \wfe \andl\\
&\gap y \in \dom(s) \andl\\
&\gap \forall j < k,\\
&\gap y \mapsto T_c \ldefs{\seq{l = v}\;\seq{m(x)=t}} \in s,\\
&\gap \Gamma \ts T \expand_y \seq{D},\\
&\gap (\forall m_i : S \rightarrow U \in \seq{D},\\
&\gap\gap t_i \in \rele j {\Gamma \envplus{x : S}} s U)\\
&\gap \andl\\
&\gap (\forall l_i : V \in \seq{D},\\
&\gap\gap v_i \in \relv j \Gamma s V)\\
&\}
\end{align*}

\subsubsection{$\mathcal{E}$}
$\rele k \Gamma s T$ defines the set of terms that appear to have type
$T$ when taking at most $k$ steps. $s$ must agree with $\Gamma$, but
$\Gamma$ can additionaly contain variables not in $s$. This is needed
for checking methods in $\mathcal{V}$ above, and for relating open
terms. If $k > 0$, $\mathcal{E}$ completes the store to agree with
$\Gamma$. It then states that if it can reduce $t$ in the completed
store to an irreducible term in $j < k$ steps, then this term must be
in a corresponding $\mathcal{V}$ set.
\begin{align*}
&\rele k \Gamma s T = \{ t \ |\\
&\gap \Gamma \ts T \wfe \andl\\
&\gap k = 0 \orl (\forall j \leq k,\\
&\gap\gap \forall s' \in \rels k \Gamma s ,\\
&\gap\gap \reductionl t {s'} {t'} {s''} j \andl\\
&\gap\gap \irred {t'} {s''} \impliesl\\
&\gap\gap \forall \Gamma' \in \relg k {s''} \Gamma ,\\
&\gap\gap {t'} \in \relv {k-j-1} {\Gamma'} {s''} T)\\
&\}
\end{align*}

\subsubsection{$\mathcal{S}$}
$\rels k \Gamma s$ for $k > 0$ defines the set of completed stores
that agree with $\Gamma$ on each variable when taking at most $k-1$
steps.
\begin{align*}
&\rels k \Gamma s = \{ s, \seq{x_i \mapsto c_i}^{m \leq i < n} \ |\\
&\gap s = \seq{x \mapsto c}^m \andl \Gamma = \seq{x : T}^n \andl \\
&\gap m\leq n \andl \forall i, m \leq i < n,\\
&\gap c_i \in \relv {k-1} \Gamma {s, \seq{x_j \mapsto c_j}^{m \leq j < i}} {T_i}\\
&\}
\end{align*}

\subsubsection{$\mathcal{G}$}
$\relg k s \Gamma$ defines a singleton set of a completed environment
that agrees with a store $s$ by simply copying the constructor type
from the store for each missig variable.
\begin{align*}
&\relg k s \Gamma = \{ \Gamma, \seq{x_i : T_{c_i}}^{m \leq i < n} \ |\\
&\gap \Gamma = \seq{x : T}^m \andl s = \seq{x \mapsto c}^n \andl\\
&\}
\end{align*}

\subsubsection{$\Gamma \ds t : T$}

$\Gamma \ds t : T$ is simplfy defined as $t \in \rele k \Gamma - T,
\forall k$.

\subsection{Proof of the Fundamental Theorem}

The fundamental theorem is the implication from $\Gamma \ts t : T$ to
$\Gamma \ds t : T$. Type safety is a straightforward corollary of this
theorem.

@@

%\appendix
%\section{Appendix Title}

\acks
\todo{acks}
